---
title: "MODEL 1 SUPPORT VECTOR MACHINE"
author: "REY P. PENDANG"
date: "2022-12-16"
output:
  pdf_document: default
  html_document: default
---

```{r}
# Load Packages

library(dplyr)    # for data wrangling
library(readr)    #Load dataset
library(ggplot2)  # for awesome graphics
library(rsample)  # for data splitting
library(caret)    # for classification and regression training
library(kernlab)  # for fitting SVMs
library(modeldata) #for Failure.binary data
library(pdp)      # for partial dependence plots, etc.
library(vip)      # for variable importance plots
library(forcats)
```

# SUPPORT VECTOR MACHINE

# Load the data set

The data frame output of data reprocessing converted into to "csv", which will be used for entire project.


```{r}
dt <- read_csv("normalRad.CSV")
View(dt)
head(dt)

# Load Failure.binary data

dt$Failure.binary=as.factor(dt$Failure.binary)
```

# CREATING THE TRAINING (80%) AND TEST (20%) SETS

```{r}
set.seed(123)  # for reproducibility

churn_split <- initial_split(dt, prop = 0.8, strata = "Failure.binary")
split_train <- training(churn_split)
split_test  <- testing(churn_split)
```


```{r}
caret::getModelInfo("svmLinear")$svmLinear$parameters # Linear (i.e., soft margin classifier)


caret::getModelInfo("svmPoly")$svmPoly$parameters # Polynomial kernel


caret::getModelInfo("svmRadial")$svmRadial$parameters # Radial basis kernel
```

# RUNNING SUPPORT VECTOR MACHINE MODEL IN TRAINING PHASE

Using **split_train**, we can tune an SVM model with radial basis kernel.

```{r}
set.seed(1854)  # for reproducibility
split_svm <- train(
  Failure.binary ~ ., 
  data = split_train,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  trControl = trainControl(method = "cv", number = 10),
  tuneLength = 10
)
```

# PLOT AND PRINT SVM MODEL WITH WITH RADIAL BASIS KERNEL.

```{r}
# Plot results
ggplot(split_svm) + theme_light()

# Print results
split_svm$results
```

# CONTROLING PARAMETER

```{r}
class.weights = c("No" = 1, "Yes" = 10)

# Control params for SVM
ctrl <- trainControl(
  method = "cv", 
  number = 10, 
  classProbs = TRUE,                 
  summaryFunction = twoClassSummary  # also needed for AUC/ROC
)

split_train$Failure.binary=fct_recode(split_train$Failure.binary,No="0",Yes="1")

```

# PRINTING THE AUC VALUES DURING TRAINING

```{r}
# Tune an SVM
set.seed(5628)  # for reproducibility
train_svm_auc <- train(
  Failure.binary ~ ., 
  data = split_train,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  metric = "ROC",  # area under ROC curve (AUC)       
  trControl = ctrl,
  tuneLength = 10
)

# Print the results
train_svm_auc$results
confusionMatrix(train_svm_auc)
```

# PRINTING THE TOP 20 IMPORTANT FEATURES DURING TRAINING

```{r}
prob_yes <- function(object, newdata) {
  predict(object, newdata = newdata, type = "prob")[, "Yes"]
}

# Variable importance plot
set.seed(2827)  # for reproducibility
vip(train_svm_auc, method = "permute", nsim = 5, train = split_train, 
    target = "Failure.binary", metric = "auc", reference_class = "Yes", 
    pred_wrapper = prob_yes)
```


# PRINTING THE AUC VALUES DURING TESTING

```{r}
split_test$Failure.binary=fct_recode(split_test$Failure.binary,No="0",Yes="1")

# Tune an SVM with radial 
set.seed(5628)  # for reproducibility
test_svm_auc <- train(
  Failure.binary ~ ., 
  data = split_test,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  metric = "ROC",  # area under ROC curve (AUC)       
  trControl = ctrl,
  tuneLength = 10
)

# Printing the results
test_svm_auc$results
confusionMatrix(test_svm_auc)
```