---
title: "MODEL 1 GRADIENT BOOSTING MODEL"
author: "REY P. PENDANG"
date: "2022-12-16"
output:
  pdf_document: default
  html_document: default
---


```{r}

# Load Packages
library(dplyr)    # for general data wrangling needs
library(tidyverse)# for filtering 
library(rsample)   # for creating validation splits
library(h2o)      # for a java-based implementation of GBM variants
library(xgboost)  # for fitting extreme gradient boosting
library(gbm)      # for original implementation of regular and stochastic GBMs
library(ROCR)
library(pROC)
library(h2o)        # a java-based implementation of random forest
h2o.init()
```

# RUNNING BASIC GBM MODEL

```{r}
set.seed(123)  # for reproducibility
dt<- read_csv("normalRad.csv")


dt$Institution=as.factor(dt$Institution)
split <- initial_split(dt, strata = "Failure.binary")
traindt <- training(split)
testdt <- testing(split)
```


```{r}
gradientBoostingModel_1 <- gbm(
  formula = Failure.binary ~ .,
  data = traindt,
  distribution = "bernoulli",  # SSE loss function
  n.trees = 5000,
  shrinkage = 0.1,
  n.minobsinnode = 10,
  cv.folds = 10

)
```

# FIND INDEX FOR NUMBER TREES WITH MINIMUM CV ERROR

```{r}
best <- which.min(gradientBoostingModel_1$cv.error)
```

# GET MSE AND COMPUTE RMSE
```{r}
sqrt(gradientBoostingModel_1$cv.error[best])
```

# PLOTTING THE ERROR CURVE
```{r}
gbm.perf(gradientBoostingModel_1, method = "cv")
```

# CREATE GRID SEARCH

```{r}
hyper_grid <- expand.grid(
  learning_rate = c(0.3, 0.1, 0.05, 0.01, 0.005),
  logloss = NA,
  trees = NA,
  time = NA
)

# EXECUTE GRID SEARCH
for(i in seq_len(nrow(hyper_grid))) {
  
  # fit gbm
  set.seed(123)  # for reproducibility
  train_time <- system.time({
    m <- gbm(
      formula = Failure.binary ~ .,
      data = traindt,
      distribution = "bernoulli",
      n.trees = 5000, 
      shrinkage = hyper_grid$learning_rate[i], 
      interaction.depth = 3, 
      n.minobsinnode = 10,
      cv.folds = 10 
    )
  })
  
  # adding SSE, trees, and training time to results
  hyper_grid$logloss[i]  <- sqrt(min(m$cv.error))
  hyper_grid$trees[i] <- which.min(m$cv.error)
  hyper_grid$Time[i]  <- train_time[["elapsed"]]
  
}
```

# RESULTS

```{r}
arrange(hyper_grid, logloss)
```


# SEARCH GRID
```{r}
hyper_grid <- expand.grid(
  n.trees = 6000,
  shrinkage = 0.01,
  interaction.depth = c(3, 5, 7),
  n.minobsinnode = c(5, 10, 15)

)
```

# CREATING THE MODEL FIT FUNCTION

```{r}
model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = Failure.binary ~ .,
    data = traindt,
    distribution = "bernoulli",
    n.trees = n.trees,
    shrinkage = shrinkage,
    interaction.depth = interaction.depth,
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))

}
```

# PERFORMING SEARCH GRID WITH FUNCTIONAL PROGRAMMING
```{r}
hyper_grid$logloss <- purrr::pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
  )
)

# RESULTS 
arrange(hyper_grid, logloss)
```

# REFINED HYPERPARAMETER GRID

```{r}
hyper_grid <- list(
  sample_rate = c(0.5, 0.75, 1),              # row subsampling
  col_sample_rate = c(0.5, 0.75, 1),          # col subsampling for each split
  col_sample_rate_per_tree = c(0.5, 0.75, 1)  # col subsampling for each tree
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "logloss",
  stopping_tolerance = 0.001,   
  stopping_rounds = 10,         
  max_runtime_secs = 60*60      
)
```

# PERFORMING GRID SEARCH

```{r}
traindt$Failure.binary=as.factor(traindt$Failure.binary)

h2o.init()

grid <- h2o.grid(
  algorithm = "gbm",
  grid_id = "gbm_grid",
  y = "Failure.binary",
  training_frame = as.h2o(traindt),
  hyper_params = hyper_grid,
  ntrees = 10,#supposedly 6000
  learn_rate = 0.01,
  max_depth = 7,
  min_rows = 5,
  nfolds = 10,
  stopping_rounds = 10,
  stopping_tolerance = 0,
  stopping_metric="logloss",
  search_criteria = search_criteria,
  seed = 123

)
```

# COLLECT THE RESULTS AND SORT BY OUR MODEL PERFORMANCE METRIC OF CHOICE

```{r}
grid_perf <- h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "logloss", 
  decreasing = FALSE
)
```


```{r}
grid_perf
```

# GRAB THE MODEL_ID FOR THE TOP MODEL, CHOSEN BY CROSS VALIDATION ERROR

```{r}
best_model_id <- grid_perf@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)

# GETTING THE PERFORMANCE METRICS ON THE BEST MODEL

h2o.performance(model = best_model, xval = TRUE)
```


```{r}
library(recipes)
xgb_prep <- recipe(Failure.binary ~ ., data = traindt) %>%
  step_integer(all_nominal()) %>%
  prep(training = traindt, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Failure.binary")])
Y <- xgb_prep$Failure.binary
Y=as.numeric(Y)-1
```


```{r}
set.seed(123)
ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 6000,
  objective = "binary:logistic",
  early_stopping_rounds = 50, 
  nfold = 10,
  params = list(
    eta = 0.1,
    max_depth = 3,
    min_child_weight = 3,
    subsample = 0.8,
    colsample_bytree = 1.0),
  verbose = 0
)

```

# MINIMUM TEST CV RMSE

```{r}
min(ames_xgb$evaluation_log$test_logloss_mean)
```


```{r}
# hyperparameter grid
hyper_grid <- expand.grid(
  eta = 0.01,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = 0.5, 
  colsample_bytree = 0.5,
  gamma = c(0, 1, 10, 100, 1000),
  lambda = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  alpha = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  logloss = 0,          # a place to dump RMSE results
  trees = 0          # a place to dump required number of trees
)

# grid search
for(i in seq_len(nrow(hyper_grid))) {
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 100,#supposedly 4000
    objective = "binary:logistic",
    early_stopping_rounds = 50, 
    nfold = 10,
    verbose = 0,
    params = list( 
      eta = hyper_grid$eta[i], 
      max_depth = hyper_grid$max_depth[i],
      min_child_weight = hyper_grid$min_child_weight[i],
      subsample = hyper_grid$subsample[i],
      colsample_bytree = hyper_grid$colsample_bytree[i],
      gamma = hyper_grid$gamma[i], 
      lambda = hyper_grid$lambda[i], 
      alpha = hyper_grid$alpha[i]
    ) 
  )
  hyper_grid$logloss[i] <- min(m$evaluation_log$test_logloss_mean)
  hyper_grid$trees[i] <- m$best_iteration
}
```


```{r}
# results
hyper_grid %>%
  filter(logloss > 0) %>%
  arrange(logloss) %>%
  glimpse()
```


```{r}
# optimal parameter list
params <- list(
  eta = 0.01,
  max_depth = 3,
  min_child_weight = 3,
  subsample = 0.5,
  colsample_bytree = 0.5
)
```


```{r}
# 
# # train final model
# traindt$Institution=fct_recode(traindt$Institution, "1" = "A", "2" ="B","3"="C","4"="D")
# traindt$Institution=as.numeric(traindt$Institution)
# traindt=as.matrix(traindt)

xgb.fit.final <- xgboost(
  params = params,
  data = X,
  label = Y,
  nrounds = 3944,
  objective = "binary:logistic",
  verbose = 0
)
```

```{r}
# Compute predicted probabilities on training data
m1_prob <- predict(xgb.fit.final, X, type = "prob")

# Compute AUC metrics for cv_model1,2 and 3 
perf1 <- prediction(m1_prob,traindt$Failure.binary) %>%
  performance(measure = "tpr", x.measure = "fpr")


# Plot ROC curves for cv_model1,2 and 3 
plot(perf1, col = "black", lty = 2)


# ROC plot for training data
roc( traindt$Failure.binary ~ m1_prob, plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, col="black", lwd=2, print.auc=TRUE)

xgb_prep <- recipe(Failure.binary ~ ., data = testdt) %>%
  step_integer(all_nominal()) %>%
  prep(training = testdt, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Failure.binary")])

# Compute predicted probabilities on training data
m2_prob <- predict(xgb.fit.final, X, type = "prob")

# Compute AUC metrics for cv_model1,2 and 3 
perf2 <- prediction(m2_prob,testdt$Failure.binary) %>%
  performance(measure = "tpr", x.measure = "fpr")


# Plot ROC curves for cv_model1,2 and 3 
plot(perf2, col = "black", lty = 2)


# ROC plot for training data
roc( testdt$Failure.binary ~ m2_prob, plot=TRUE, legacy.axes=FALSE, 
    percent=TRUE, col="black", lwd=2, print.auc=TRUE)

```



```{r}
# variable importance plot
vip::vip(xgb.fit.final,num_features=20) 
```


